---
title: "srs_PCA"
author: "Christen Bossu"
date: "06/06/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In terminal on your cluster, you'll need to filter the merged gatk vcf file. If the vcf file doesn't have the MISSING info field (mine doesn't), you can filter in vcftools

vcftools option below: 

If vcftools is a module that your cluster has installed globablly, load it. Otherwise, you can install it yourself and put it in your path and will not need the `module load vcftools` line.

```
module load vcftools
vcftools --vcf COYEP1-4c.merged_gatk.srs_filt.vcf --out COYEP1-4c.merged_gatk.srs_filt.75. --recode --maf 0.05 --max-maf 0.95 --min-alleles 2 --max-alleles 2 --max-missing 0.5 
```

Get allele depths for your filtered vcf file. Here, because that's too large, the trial file will allow us to get 50,000 sites of the 15464410 sites using bcftools. For all data, just remove the 

`head -500001` 

line

```
bcftools query -f '%CHROM\t%POS[\t%AD]\n' COYEP1-4c.merged_gatk.srs_filt.75.vcf|head -50001 > COYEP1_P4c.allele_depths.subset.txt
```

## Covariance analysis
First install and load packages you'll need
```{r}
#remotes::install_github("eriqande/srsStuff", upgrade = FALSE)
library(srsStuff)
library(tidyverse)
library(gdata)
library(RColorBrewer)

```

Next specify the files with the individual names and the allele depth file
```{r}
snames <- read_lines("data/srs/test.ind")

AllCOYE<-srs_covar(file="data/srs/COYEP1_4c.allele_depths.subset.txt", sample_names=snames, freq_thresh = 0)
saveRDS(AllCOYE, "data/srs/COYE.srs_filtered.202ind.srs.subset.rds")
AllCOYE <- readRDS("data/srs/COYE.srs_filtered.202ind.srs.subset.rds")
```

Incorporate the meta data
```{r}
meta<-read.xls("data/srs/COYE_master_Updated06062022.xlsx",sheet=3) %>% dplyr::select(Sample_data,State,Location) %>% rename(vcf_name=Sample_data,Region_site=State,Region_pop=Location) %>% filter(vcf_name %in% AllCOYE$sample_names) %>% distinct() %>% filter(Region_site!="NA") %>% arrange(Region_site,vcf_name)
dim(meta)
meta %>% distinct(Region_site)
```

Next to create the covariance matrix.

# A function from Eric to get everything prepped and tidy
```{r}
prep_output <- function(A) {

  # first for each individual i, get the number of sites at which
  # that individual had a read and also another individual, j, averaged
  # over all j.  This tells us on average how many markers went into
  # compute the covariance for each individual.
  m <- A$M
  diag(m) <- 0
  ave_sites <- tibble(
    vcf_name = A$sample_names,
    ave_sites = rowMeans(m) * ncol(m) / (ncol(m) - 1)
  )

  # do the whole eigendecomposition on the standard covariance matrix.
  eig <- eigen(A$Cov)
  colnames(eig$vectors) <- sprintf("PC-%02d", 1:ncol(eig$vectors))


  pca_tib <- as_tibble(eig$vectors[,1:6]) %>%
    mutate(vcf_name = A$sample_names) %>%
    dplyr::select(vcf_name, everything())

  pca_long <- pca_tib %>%
    tidyr::gather(., key = "PC", "val", -vcf_name)

  # then expand a grid of the possible comparisons (ordered)
  pca_pairs <- expand.grid(vcf_name = pca_tib$vcf_name,
                           PCx = sprintf("PC-%02d", 1:6),
                           PCy = sprintf("PC-%02d", 1:6),
                           stringsAsFactors = FALSE) %>%
    tibble::as_tibble() %>%
    dplyr::left_join(., pca_long, by = c("vcf_name", "PCx" = "PC")) %>%
    dplyr::rename(val_x = val) %>%
    dplyr::left_join(pca_long, by = c("vcf_name", "PCy" = "PC")) %>%
    dplyr::rename(val_y = val)  %>%
    left_join(ave_sites, by = "vcf_name") %>%
    left_join(meta, by = "vcf_name")

  WaG <- (A$IBS - A$Mt_S) / (1 - A$Mt_S)
  eig <- eigen(WaG)
  colnames(eig$vectors) <- sprintf("PC-%02d", 1:ncol(eig$vectors))


  pca_tib <- as_tibble(eig$vectors[,1:6]) %>%
    mutate(vcf_name = A$sample_names) %>%
    dplyr::select(vcf_name, everything())

  pca_long <- pca_tib %>%
    tidyr::gather(., key = "PC", "val", -vcf_name)

  pcp <- expand.grid(vcf_name = pca_tib$vcf_name,
                     PCx = sprintf("PC-%02d", 1:6),
                     PCy = sprintf("PC-%02d", 1:6),
                     stringsAsFactors = FALSE) %>%
    tibble::as_tibble() %>%
    dplyr::left_join(., pca_long, by = c("vcf_name", "PCx" = "PC")) %>%
    dplyr::rename(WaG_x = val) %>%
    dplyr::left_join(pca_long, by = c("vcf_name", "PCy" = "PC")) %>%
    dplyr::rename(WaG_y = val)

  # then join those both and return them
  left_join(pcp, pca_pairs,
            by = c("vcf_name", "PCx", "PCy"))
}
```

##COYE Breeding Site Analysis
This puts the sites in order and allows you the pick a color for the groupings using RColorBrewer palettes.
```{r}
breedingCOYE.prep <- prep_output(AllCOYE) #%>%
  #filter(Region_pop == "Breeding")

# change the factor levels of sites!
factor.levels <- c("CA", "MT","WA","BC", "AK",
                   "WY", "NV","NM",
                   "NY", "PA", "NB","QUE","NC",
                   "KY", "MI","ON","AB","KS","AL",
                   "AZ","FL")
                   
breedingCOYE.prep$Region_site <- factor(x = breedingCOYE.prep$Region_site,
                                   levels = factor.levels)


West.cols <- brewer.pal(n = 6, name = "PuRd")[c(1:5)]
SW.cols <- brewer.pal(n = 9, name = "BuGn")[c(3:5)]
NE.cols <- brewer.pal(n = 9, name = "Blues")[c(2:7)]
MW.cols <- brewer.pal(n = 9, name = "YlOrRd")[c(2:8)]
Res.cols<-brewer.pal(n = 9, name = "Set1")[c(8,9)]
pca.color <- c(West.cols,SW.cols,NE.cols,MW.cols,Res.cols)
```

##Initial Plots
Facet grid:
```{r}
bp <- ggplot(breedingCOYE.prep, aes(x = val_x, y = val_y, color = Region_site)) +
  facet_grid(PCy ~ PCx) +
  scale_color_manual(values = pca.color) +
  geom_point(size = 3, stroke = 0.15) +
  theme_bw() + theme(legend.key.size = unit(0.1, "cm")) 
  
bp
```

##Figure out which indivduals to remove

```{r}
cov_and_M <- function(A) {
  tibble(vcf_name_1 = rep(A$sample_names, times = nrow(A$Cov)),
         vcf_name_2 = rep(A$sample_names, each = ncol(A$Cov)),
         covar = as.vector(A$Cov),
         M = as.vector(A$M),
         WaG = as.vector((A$IBS - A$Mt_S) / (1 - A$Mt_S))
  )
}

CM <- cov_and_M(AllCOYE)
#CM %>% filter(vcf_name_1=="11N3564")
#CM %>%
#   filter(vcf_name_1 < vcf_name_2) %>%
#   arrange(desc(M)) %>%
#   ggplot(aes(x = covar, y = 2, colour = log10(M))) +
#   geom_jitter(width = 0, height = 1, alpha = 0.6) +
#   scale_colour_viridis_c()


CM_cut <- CM %>%
  filter(vcf_name_1 < vcf_name_2) %>%
  arrange(desc(M)) %>%
  mutate(m_group = cut(M, breaks = c(0, 1e6, 5e6, 1e7, 1e8, 2e8)))


g <- ggplot(CM_cut, aes(x = covar, y = 2, colour = log10(M), size = 2)) +
  geom_jitter(width = 0, height = 1) +
  scale_colour_viridis_c() +
  facet_wrap(~m_group, ncol = 1) +
  theme_bw()

g
```

Save file.
```{r}
ggsave(g, filename = "plots/srs_plots/coye-50missing-covar-and-M-facets.pdf",
      width = 20, height = 20)
```

##Create table to individuals to remove.
First, choose the covariance cut off that works for your data. Then look at the clustering, does it make sense (are indvidiuals with high covariance from the same population?).

```{r}
outlier_CM_cut <- CM_cut %>% 
  filter(covar > 0.4)
outlier_covar <- unique(c(unique(outlier_CM_cut$vcf_name_1), unique(outlier_CM_cut$vcf_name_2)))


meta[meta$vcf_name %in% outlier_covar,]
```

```{r}
meta[meta$vcf_name %in% outlier_covar,] %>% write.table("output_files/srs_results/COYEind_outlier.cov_gr_.4.txt",row.names=F,quote=F,sep="\t")
```

Need to install packages for get gRbase to run.
```{r}
#if (!requireNamespace("BiocManager", quietly = TRUE))
#    install.packages("BiocManager")

#BiocManager::install("graph")
#BiocManager::install("Rgraphviz")
#BiocManager::install("RBGL")
```

Remove just the very related first (cov > 0.4), #then increase cov removal (cov > .3)


```{r}
library(graph)
library(Rgraphviz)
library(RBGL)
library(gRbase)
```

if libraries above installed, skip to this step to identify which individual from cluster to keep!
```{r}
matchers <- outlier_CM_cut %>%
  dplyr::select(vcf_name_1, vcf_name_2) %>%
  rename(indiv_1 = vcf_name_1, indiv_2 = vcf_name_2)
  

edgelist <- lapply(1:nrow(matchers), function(x) c(matchers$indiv_1[x], matchers$indiv_2[x]))
gR <- gRbase::ug(edgelist)
CnC <- graph::connComp(gR)
CnC_tib <- tibble(
  indiv = unlist(CnC),
  cluster = rep(1:length(CnC), sapply(CnC, length))
) 
CnC_tib %>% rename(vcf_name=indiv) %>% left_join(meta[meta$vcf_name %in% outlier_covar,]) %>% write.table("output_files/srs_results/COYE.srs.cov.4.Cluster_choose1.txt",row.names=F,quote=F,sep="\t")
CnC_tib
```

Individuals to remove. Here, I choose one individual per cluster to keep (the rest are removed). 

```{r}
#Keep these for cov group > 0.4
keep.these <- c("MOR_12",
                   "17N02622",
                   "16N0010",
                   "02N8421",
                   "16N0007",
                   "16N0012",
                   "98N2078")

#Keep these for cov group > 0.27
keep.these <- c("MOR_12",
                   "17N02622",
                   "16N0005",
                   "16N0007",
                   "98N2078",
                   "02N8421")


remove.these <- CnC_tib %>%
  dplyr::select(indiv) %>%
  filter(!indiv %in% keep.these) %>%
  pull()
remove.these %>% write.table("output_files/srs_results/COYE.remove_indiv.cov.4.txt",row.names=F,quote=F,sep="\t")
remove.these.snames.index <- c(1:202)[snames %in% remove.these]
```

Re-do the eigenvalues and PCA. The creates factor levels and colors for each population.

```{r}
AllCOYE.new <- AllCOYE

AllCOYE.new$IBS <- AllCOYE$IBS[-c(remove.these.snames.index),
                          -c(remove.these.snames.index)]
AllCOYE.new$Cov <- AllCOYE$Cov[-c(remove.these.snames.index),
                          -c(remove.these.snames.index)]
AllCOYE.new$M <- AllCOYE$M[-c(remove.these.snames.index),
                          -c(remove.these.snames.index)]
AllCOYE.new$Mt_S <- AllCOYE$Mt_S[-c(remove.these.snames.index),
                          -c(remove.these.snames.index)]
AllCOYE.new$sample_names <- snames[!snames %in% remove.these]

breedingCOYE.new.prep <- prep_output(AllCOYE.new) 

# change the factor levels of sites!
factor.levels <- c("CA", "MT","WA","BC", "AK","AB",
                   "WY", "NV","NM",
                   "NY", "PA", "NB","QUE",
                   "KY", "MI","ON","KS",
                   "FL","AL","NC",
                   "AZ")
                   
breedingCOYE.new.prep$Region_site <- factor(x = breedingCOYE.new.prep$Region_site,
                                   levels = factor.levels)

West.cols <- brewer.pal(n = 6, name = "PuRd")[c(1:6)]
SW.cols <- brewer.pal(n = 9, name = "BuGn")[c(3:5)]
NE.cols <- brewer.pal(n = 9, name = "Blues")[c(3:6)]
MW.cols <- brewer.pal(n = 9, name = "YlOrRd")[c(2:5)]
South.cols<-brewer.pal(n = 9, name = "Greys")[c(3:5)]
Res.cols<-brewer.pal(n = 9, name = "YlOrRd")[c(7)]
pca.color <- c(West.cols,SW.cols,NE.cols,MW.cols,South.cols,Res.cols)

bp <- ggplot(breedingCOYE.new.prep, aes(x = val_x, y = val_y, color = Region_site)) +
  facet_grid(PCy ~ PCx) +
  scale_color_manual(values = pca.color) +
  geom_point(size = 3, stroke = 0.15) +
  theme_bw() + theme(legend.key.size = unit(0.1, "cm"))
bp
```

PCs 1 and 2 with > 0.2 removed. Fingers crossed here! Does it change?
```{r}
pc.eigenvalues <- eigen(AllCOYE.new$Cov)$values
pc.var.expl <- pc.eigenvalues[1:2] / sum(pc.eigenvalues) * 100

pc.var.expl

PC12.prep <- breedingCOYE.new.prep %>%
  filter(PCx == "PC-01", PCy == "PC-02")

pca12 <- ggplot(PC12.prep, aes(x = val_x, y = val_y, color = Region_site)) +
  scale_color_manual(values = pca.color) +
  geom_point(size = 3, stroke = 0.15) +
  #geom_text(aes(x = val_x, y = val_y,label=Region_site))+
  theme_bw() + theme(aspect.ratio=1) + 
  theme(legend.key.size = unit(0.1, "cm"))+
  xlab("PC1: 14.978055%") +
  ylab("PC2: 0.715093%") +
  guides(fill = guide_legend(override.aes=list(shape=22, stroke = 0.1, size = 6)),
         shape = guide_legend(override.aes=list(stroke = 0.35, size = 4)))
  
pca12
```


```{r}
ggsave(pca12, filename = "plots/srs_plots/COYE.PC1_PC2.cov_gr_.4.195ind.pdf",
      width = 7, height = 7)
```

Let's look at PC2 and PC3
```{r}
pc.eigenvalues <- eigen(AllCOYE.new$Cov)$values
pc.var.expl <- pc.eigenvalues[2:3] / sum(pc.eigenvalues) * 100
pc.var.expl
PC23.prep <- breedingCOYE.new.prep %>%
  filter(PCx == "PC-02", PCy == "PC-03")

pca23 <- ggplot(PC23.prep, aes(x = val_x, y = val_y, color = Region_site)) +
  scale_color_manual(values = pca.color) +
  geom_point(size = 3, stroke = 0.15) +
  theme_bw() +
  theme(legend.key.size = unit(0.1, "cm")) +
  xlab("PC1: 0.715093%") +
  ylab("PC2: 0.680002%") 
  
pca23
```

```{r}
ggsave(pca23, filename = "plots/srs_plots/COYE.PC2_PC3.cov_gr_.4.pdf",
      width = 7, height = 7)
```

