---
title: "ngsRelate"
author: "Christen Bossu"
date: "6/6/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Motivation behind testing relatedness

One reason you may have high genetic covariance among individuals (identified in the previous srs section) is relatedness. Parent/offpsring relationships and full siblings are more genetically similar and therefore have higher genetic covariance. To test relatedness among the indivdiuals, and to potentially confirm the removal of those indivdials from subsequent analyses, I use ngsRelate in the terminal.

Here, define the path to your ngsRelate software program, which you can git here:

https://github.com/ANGSD/NgsRelate


```
NGSRELATE="/ocean/projects/bio210095p/cbossu/programs/ngsRelate/ngsRelate"
```

You can use this on a glf file from ANGSD or the vcf file you used in the previous analysis. These have very different codes, and the vcf one is below.

By default, NgsRelate will estimate the allele frequencies using the individuals provided in the VCF files.

```
$NGSRELATE/ngsRelate -h data/ngsRelate_data/COYEP1-4c.merged_gatk.srs_filt.subset.recode.vcf -z data/test.ind -O output_files/ngsRelate_results/COYEP1-4c.merged_gatk.srs_filt.subset.recode.vcf.ngsRelate.res -T GT -c 1
```

Now it's time to move to R. Load the libraries, we'll need
```{r}
library(tidyverse)
library(gdata)
library(ggpubr)
```


##Related plots
So you will need to do some adjusting of the file to then get it into R for plotting. Basically you added the the individual with teh -z function above, but now you need to add the region/location for each individual.

```{r}
meta<-read.xls("data/srs/COYE_master_Updated06062022.xlsx",sheet=3) %>% dplyr::select(Sample_data,State) %>% rename(Sample=Sample_data,Region_site=State)

relSC<-read_delim("output/ngsRelate_results/COYEP1-4c.merged_gatk.srs_filt.subset.recode.vcf.ngsRelate.res",delim="\t") %>%  
  rename(IndA=ida,IndB=idb) %>% 
  dplyr::select(IndA,IndB,everything())

RegA<-relSC %>% dplyr::select(IndA) %>% rename(Sample=IndA) %>% left_join(meta) %>% rename(RegionA=Region_site,IndA=Sample)
RegB<-relSC %>% dplyr::select(IndB) %>% rename(Sample=IndB) %>% left_join(meta) %>% rename(RegionB=Region_site,IndB=Sample)

relSC_dat<-relSC %>% left_join(RegA) %>% left_join(RegB) %>% 
  unite("Pair",IndA:IndB,sep="_",remove=F) %>% 
  rename(K0=J9,K1=J8,K2=J7) %>% 
  mutate(RegionBoth=if_else(RegionA==RegionB,RegionA,"mixture")) %>%
  dplyr::select(-X12,-X13,-a,-b) %>% filter(RegionBoth!="mixture")
relSC %>% distinct(RegionBoth)
```

```{r}
library(ggpubr)
g<-relSC_dat  %>% 
  ggplot(aes(K0,K1,col=RegionBoth))+geom_point(size=2,alpha=0.8) +scale_colour_manual(values=colors) + theme_bw() +theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + geom_text(aes(label=Pair),hjust=-0.1, vjust=0,size=2) + theme(aspect.ratio = 1)

pdf("COYE.NGSrelate.k0_k1.plot.RegionA.labelPair.pdf",useDingbats = F)
ggarrange(g,
          labels = c("A.  "),
          ncol = 2, nrow = 2)
dev.off()

```