---
title: "Admixture"
author: "Christen Bossu"
date: "6/6/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = 'output/admixture_results')
```

## Format input file
From your vcf file, first remove your related indivdiuals. 

```
vcftools --vcf data/ngsRelate/COYEP1-4c.merged_gatk.srs_filt.subset.recode.vcf --out data/ngsRelate/coye.rm_rel.subset --remove srs_results/COYE.remove_indiv.cov.4.txt  --recode
```
Then create a ped/map file.
```
plink --vcf data/ngsRelate/coye.rm_rel.subset.recode.vcf --out data/admixture/coye.subset --aec --recode
```
The map file needs some formating here.

For some reason, admixture doesn't like have alternative chromosomes, so for each ped file, I humanize it, either removing the chromosome/scaffold names that admixture doesn't like and replacing with 0 or 1.

```
cd data/admixture
cat coye.subset.map|awk '{print"1\t"$1"."$4"\t"$3"\t"$4}' > coye.subset.no_chr.map
```

Now run plink so you have the same prefix for map and ped files

```
plink --ped coye.subset.ped --map coye.subset.no_chr.map --out coye.subset.no_chr --make-bed --aec
```

## Running admixture is relatively easy
I have admixture in my path, so make sure you do too.

```
for j in 1 2 3 ;
do
for K in 2 3 4 5 6 7 8; \
do admixture --cv coye.subset.no_chr.bed $K | tee log_coye${K}.run"$j".out;
mv coye.subset.no_chr.${K}.Q ../../output/admixture_results/coye.subset.no_chr.run"$j".${K}.Q;
mv coye.subset.no_chr.${K}.P ../../output/admixture_results/coye.subset.no_chr.run"$j".${K}.P;
done
done
```

Then for admixture we look at the cv error, where the number in parentheses is the standard error of the cross-validation error estimate.

We can easily plot these values for comparison, which makes it fairly clear that K = 3 is a sensible modeling choice.

I clean up these log files to read into R
```
grep -h CV log.coye*.out|sed 's/(K=//g'|sed 's/)://g'|cut -f 3- -d " " > COYE.cv_error.txt
```

```{r}
cv<-read_delim("../../data/admixture/COYE.cv_error.txt",delim=" ",col_names = F) %>% rename(K=X1,cv_error=X2)
cv
```


