---
title: "Ebird Range"
author: "Christen Bossu"
date: "3/8/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Ebird

Can we create a new range map from ebird, to be split into bredding genetic cluster polygons later
```{r cars}
# install.packages("remotes")
#remotes::install_github("CornellLabofOrnithology/ebirdst")
library(ebirdst)
library(viridis)
library(raster)
library(sf)
library(smoothr)
library(rnaturalearth)
library(dplyr)
library(tidyr)
library(stringr)
library(ggplot2)
# resolve namespace conflicts
select <- dplyr::select
```

## Abundance maps
First, get a key from ebird. If you already have it skip this step. If you don't, get the key, then restart R to read the key

#set access key once, until you need to get a new one
```{r}

ebirdst::set_ebirdst_access_key("XXXX",overwrite = T)
```
Download to a temp directory for the vigette, but in practice, change to permanent directory the status and trends downloads.

Use the 6 letter code for the species you are working on
```{r}
sp_path <- ebirdst_download(species = "comyel")
# load the abundance data
# this automaticaaly labels layers with their dates
abd <- load_raster("abundance", path = sp_path)
```

```{r}
mollweide <- "+proj=moll +lon_0=-90 +x_0=0 +y_0=0 +ellps=WGS84"
ne_scale <- 50

# land polygon
ne_land <- ne_countries(scale = ne_scale, returnclass = "sf") %>%
  filter(continent %in% c("North America", "South America")) %>%
  st_set_precision(1e6) %>%
  st_union() %>% 
  st_geometry()
# function to subset other features to those  within this land area
wh_subset <- function(x) {
  in_wh <- as.logical(st_intersects(x, ne_land, sparse = FALSE))
  st_transform(x[in_wh], crs = mollweide)
}
mollweide
##There is a way to download, but you don't want to do this each time. I have these on my computer, but they are also up on Gdrive.

# country lines
ne_country_lines <- ne_download(scale = ne_scale, category = "cultural",
                                type = "admin_0_boundary_lines_land",
                                returnclass = "sf") %>%
  st_geometry() %>%
  wh_subset()

ne_country_lines <- st_read("~/Dropbox/BGP/genoscape_maps/shapefiles/ne_shapefiles/ne_50m_admin_0_boundary_lines_land.shp") %>%
  st_geometry() %>%
  wh_subset()

# state lines
sf_use_s2(FALSE)
# state lines
ne_state_lines <- st_read("~/Dropbox/BGP/genoscape_maps/shapefiles/ne_shapefiles/ne_50m_admin_1_states_provinces.shp") %>%
  st_geometry() %>%
  wh_subset()

# rivers
ne_rivers <- st_read("~/Dropbox/BGP/genoscape_maps/shapefiles/ne_shapefiles/ne_50m_rivers_lake_centerlines.shp") %>%
  st_geometry() %>%
  wh_subset()

# lakes
ne_lakes <- st_read("~/Dropbox/BGP/genoscape_maps/shapefiles/ne_shapefiles/ne_50m_lakes.shp") %>%
  st_geometry() %>%
  wh_subset()

\ne_land <- st_transform(ne_land, crs = mollweide)
```
##Seasonal 

```{r}
# subset to the yellow-bellied sapsucker season definitions
comyel_dates <- filter(ebirdst_runs, species_code == "comyel") %>% 
  # just keep the seasonal definition columns
  select(setdiff(matches("(start)|(end)"), matches("year_round"))) %>% 
  # transpose
  gather("label", "date") %>% 
  # spread data so start and end dates are in separate columns
  separate(label, c("season", "start_end"), "_(?=s|e)") %>% 
  spread(start_end, date) %>% 
  select(season, start_dt, end_dt)
# did the season pass review
comyel_dates2 <- mutate(comyel_dates, pass = !(is.na(start_dt) | is.na(end_dt)))
comyel_dates2
```

```{r}
# dates for each abundance layer
weeks <- parse_raster_dates(abd)
# assign to seasons
weeks_season <- rep(NA_character_, length(weeks))
for (i in seq_len(nrow(comyel_dates2))) {
  s <- comyel_dates2[i, ]
  # skip seasona assignment if season failed
  if (!s$pass) {
    next()
  }
  # handle seasons cross jan 1 separately
  if (s$start_dt <= s$end_dt) {
    in_season <- weeks >= s$start_dt & weeks <= s$end_dt
  } else {
    in_season <- weeks >= s$start_dt | weeks <= s$end_dt
  }
  weeks_season[in_season] <- s$season
}
table(weeks_season)
```
```{r}
# drop weeks not assigned to season
week_pass <- !is.na(weeks_season)
abd <- abd[[which(week_pass)]]
weeks <- weeks[week_pass]
weeks_season <- weeks_season[week_pass]
# average over weeks in season
mean_season <- function(s) {
  calc(abd[[which(weeks_season == s)]], mean, na.rm = TRUE)
}
seasons <- unique(weeks_season)
abd_season <- lapply(seasons, mean_season) %>% 
  stack() %>% 
  setNames(seasons)
abd_season
```
This code chunk doesn't work for some reason.
```
migration_threshold <- 0.4
mig_seasons <- c("prebreeding_migration", "postbreeding_migration")
if (all(mig_seasons %in% names(abd_season))) {
  # identify areas with abundance in only one season
  abd_nz <- abd_season[[mig_seasons]] > 0
  just_pre <- mask(abd_nz[["prebreeding_migration"]],
                   abd_nz[["postbreeding_migration"]], 
                   maskvalue = 1)
  just_post <- mask(abd_nz[["postbreeding_migration"]],
                   abd_nz[["prebreeding_migration"]], 
                   maskvalue = 1)
  # count the number of cells with abundance in only one season
  n_just <- cellStats(stack(just_pre, just_post), sum)
  n_all <- cellStats(abd_nz, sum)
  # is the proportion of one season cells above the 40% threshold
  split_migration <- max(n_just / n_all, na.rm = TRUE) >= migration_threshold
} else {
  split_migration <- FALSE
}
n_just / n_all
#>  prebreeding_migration postbreeding_migration 
#>           0.0002903095           0.0000000000
```

But I don't think it's entirely necessary to split migration into spring or fall for our questions. IF you do, change this to TRUE.

```{r}
split_migration <- FALSE
#> [1] FALSE
```


```{r}
threshold_yearround <- 0.01
# decide whether to show year-round layer
if (nlayers(abd_season) == 4) {
  # annual abundance
  abd_yr <- calc(abd, fun = mean, na.rm = TRUE)
  # mask out cells that aren't occupied year-round
  year_round <- calc(abd_season > 0, fun = sum, na.rm = TRUE) == 4
  abd_yr_mask <- mask(abd_yr, year_round, maskvalue = 0)
  # determine proportion of celss that are occupied year round
  n_yr <- cellStats(abd_yr_mask > 0, sum)
  n_an <- cellStats(abd_yr > 0, sum)
  # only show year round abundance if it's above 1% of range threshold
  show_yearround <- ((n_yr / n_an) >= threshold_yearround)
} else {
  show_yearround <- FALSE
}
show_yearround

```

```{r}

bin_breaks <- calc_bins(abd_season)
```


```{r}
# project the abundance data to mollweide
# use nearest neighbour resampling to preserve true zeros
abd_season_proj <- projectRaster(abd_season, crs = mollweide, method = "ngb")
# determine spatial extent for plotting
ext <- calc_full_extent(abd_season_proj)
# set the plotting order of the seasons
season_order <- c("postbreeding_migration", "prebreeding_migration", 
                  "nonbreeding", "breeding")

# prediction region, cells with predicted value in at least one week
pred_region <- calc(abd_season_proj, mean, na.rm = TRUE)
# mask to land area
ne_land_buffer <- st_buffer(ne_land, dist = max(res(pred_region)) / 2)
pred_region <- mask(pred_region, as_Spatial(ne_land_buffer))

# remove zeros from abundnace layers
abd_no_zero <- subs(abd_season_proj, data.frame(from = 0, to = NA), 
                    subsWithNA = FALSE)

# set up plot area
par(mar = c(0 , 0, 0, 0))
plot(ne_land, col = "#eeeeee", border = NA, 
     xlim = c(ext@xmin, ext@xmax),
     ylim = c(ext@ymin, ext@ymax))
# prediction region and explicit zeros
plot(pred_region, col = "#dddddd", maxpixels = raster::ncell(pred_region),
     legend = FALSE, add = TRUE)
# lakes
plot(ne_lakes, col = "#ffffff", border =  "#444444", lwd = 0.5, add = TRUE)
# land border
plot(ne_land, col = NA, border = "#444444", lwd = 0.5, add = TRUE)
# seasonal layer
plot_seasons <- intersect(season_order, names(abd_no_zero))
for (s in plot_seasons) {
  # handle splitting of migration seasons into different colors
  if (!split_migration && s %in% c("prebreeding_migration", 
                                   "postbreeding_migration")) {
    pal_season <- "migration"
    
  } else {
    pal_season <- s
  }
  pal <- abundance_palette(length(bin_breaks$bins) - 1, pal_season)
  plot(abd_no_zero[[s]], col = pal, breaks = bin_breaks$bins,
       maxpixels = ncell(abd_no_zero[[s]]),
       legend = FALSE, add = TRUE)
}
# year round
if (show_yearround) {
  year_round_proj <- projectRaster(year_round, crs = mollweide, method = "ngb")
  plot(year_round_proj, 
       col = abundance_palette(length(bin_breaks$bins) - 1, "year_round"), 
       breaks = bin_breaks$bins,
       maxpixels = ncell(year_round_proj),
       legend = FALSE, add = TRUE)
}
# linework
plot(ne_rivers, col = "#ffffff", lwd = 0.75, add = TRUE)
plot(ne_state_lines, col = "#ffffff", lwd = 1.5, add = TRUE)
plot(ne_country_lines, col = "#ffffff", lwd = 2, add = TRUE)

# legends
legend_seasons <- plot_seasons
if (split_migration) {
  legend_seasons[legend_seasons %in% c("prebreeding_migration", 
                                       "postbreeding_migration")] <- "migration"
  legend_seasons <- unique(legend_seasons)
}
if (show_yearround) {
  legend_seasons <- c(legend_seasons, "year_round")
}
# thin out labels
lbl_at <- bin_breaks$bins^bin_breaks$power
lbl_at <- c(min(lbl_at), median(lbl_at), max(lbl_at))
lbl <- lbl_at^(1 / bin_breaks$power)
lbl <- format(round(lbl, 2), nsmall = 2)
# plot legends
for (i in seq_along(legend_seasons)) {
  pal <- abundance_palette(length(bin_breaks$bins) - 1, legend_seasons[i])
  if (i == 1) {
    axis_args <- list(at = lbl_at, labels = lbl, line = -1,
                      cex.axis = 0.75, lwd = 0)
  } else {
    axis_args <- list(at = lbl_at, labels = rep("", 3),
                      cex.axis = 0.75, lwd = 0)
  }
  legend_title <- legend_seasons[i] %>% 
    str_replace_all("_", " ") %>% 
    str_to_title()
  fields::image.plot(zlim = range(bin_breaks$bins^bin_breaks$power), 
                   legend.only = TRUE, 
                   breaks = bin_breaks$bins^bin_breaks$power, col = pal,
                   smallplot = c(0.05, 0.35, 0.01 + 0.06 * i, 0.03 + 0.06 * i),
                   horizontal = TRUE,
                   axis.args = axis_args,
                   legend.args = list(text = legend_title, side = 3, 
                                      cex = 0.9, col = "black", line = 0.1))
}
title("Common Yellowthroat Relative Abundance", 
      line = -1, cex.main = 1)
```

Range Maps
The eBird Status and Trends range maps delineate the boundary of regions with non-zero relative abundance for a given species. We’ll start by aggregating the raster layers to a coarser resolution to speed up processing, then convert the boundaries of non-zero abundance regions to polygons. We’ll also convert the prediction areas to polygons so we can distinguish where the species is predicted to not occur from where it was not modeled.

```{r}
# aggregate
abd_season_agg <- aggregate(abd_season_proj, fact = 3)
# raster to polygon, one season at a time
range <- list()
pred_area <- list()
for (s in names(abd_season_agg)) {
  # range
  range[[s]] <- rasterToPolygons(abd_season_agg[[s]], 
                                 fun = function(y) {y > 0}, 
                                 digits = 6) %>% 
    st_as_sfc() %>% 
    # combine polygon pieces into a single multipolygon
    st_set_precision(1e6) %>% 
    st_union() %>% 
    st_sf() %>% 
    # tag layers with season
    mutate(season = s, layer = "range")
  # prediction area
  pred_area[[s]] <- rasterToPolygons(abd_season_agg[[s]], 
                                     fun = function(y) {!is.na(y)}, 
                                     digits = 6) %>% 
    st_as_sfc() %>% 
    # combine polygon pieces into a single multipolygon
    st_set_precision(1e6) %>% 
    st_union() %>% 
    st_sf() %>% 
    # tag layers with season
    mutate(season = s, layer = "prediction_area")
}

# combine the sf objects for all seasons
range <- rbind(do.call(rbind, range), do.call(rbind, pred_area))
row.names(range) <- NULL
print(range)
```

```{r}
cell_area <- (1.5 * prod(res(abd_season_agg)))
range_smooth <- range %>% 
  # drop fragment polygons smaller than 1.5 times the aggregated cell size
  drop_crumbs(threshold = cell_area) %>% 
  # drop holes in polygons smaller than 1.5 times the aggregated cell size
  fill_holes(threshold = cell_area) %>% 
  # smooth the polygon edges
  smooth(method = "ksmooth", smoothness = 2)
# clip zeros to land border, range to buffered land to handle coastal species
range_split <- split(range_smooth, range_smooth$layer)
class(range_split)
class(range_smooth)
range_smooth <- rbind(
  st_intersection(range_split$range, ne_land_buffer),
  st_intersection(range_split$prediction_area, ne_land))

```

```{r}
# range map color palette
range_palette <- c(nonbreeding = "#1d6996",
                   prebreeding_migration = "#73af48",
                   breeding = "#cc503e",
                   postbreeding_migration = "#edad08",
                   migration = "#edad08",
                   year_round = "#6f4070")

# set up plot area
pdf("COYE.Ebird_range.map.pdf")
par(mar = c(0 , 0, 0, 0))
plot(ne_land, col = "#eeeeee", border = NA, 
     xlim = c(ext@xmin, ext@xmax),
     ylim = c(ext@ymin, ext@ymax))
# prediction region and explicit zeros
annual_pred_area <- filter(range_smooth, layer == "prediction_area") %>% 
  st_union()
plot(annual_pred_area, col = "#dddddd", border = NA, add = TRUE)
# lakes
plot(ne_lakes, col = "#ffffff", border =  "#444444", lwd = 0.5, add = TRUE)
# land border
plot(ne_land, col = NA, border = "#444444", lwd = 0.5, add = TRUE)
# seasonal layer
for (s in intersect(season_order, unique(range_smooth$season))) {
  # handle splitting of migration seasons into different colors
  if (!split_migration && s %in% c("prebreeding_migration", 
                                   "postbreeding_migration")) {
    col_season <- "migration"
  } else {
    col_season <- s
  }
  rng_season <- filter(range_smooth, season == s, layer == "range") %>% 
    st_geometry()
  plot(rng_season, col = range_palette[col_season], border = NA, add = TRUE)
}
# year round
if (show_yearround) {
  # find common area between all seasons
  range_combined <- filter(range_smooth, layer == "range")
  range_yearround <- range_combined[1, ]
  range_combined <- sf::st_geometry(range_combined)
  for (i in 2:length(range_combined)) {
    range_yearround <- sf::st_intersection(range_yearround, range_combined[i])
  }
  plot(st_geometry(range_yearround), 
       col = range_palette["year_round"], border = NA, 
       add = TRUE)
}
# linework
plot(ne_rivers, col = "#ffffff", lwd = 0.75, add = TRUE)
plot(ne_state_lines, col = "#ffffff", lwd = 1.5, add = TRUE)
plot(ne_country_lines, col = "#ffffff", lwd = 2, add = TRUE)

# legend
rng_legend <- rev(range_palette[legend_seasons])
names(rng_legend) <- names(rng_legend) %>% 
    str_replace_all("_", " ") %>% 
    str_to_title()
legend("bottomleft", legend = names(rng_legend), fill = rng_legend)
title("Common Yellowthroat Seasonal Range Map", 
      line = -1, cex.main = 1)
dev.off()
```

```{r}
class(range_smooth)
breed_season <- filter(range_smooth, season == "breeding", layer == "range") #%>% 
    #st_geometry()
class(breed_season)

breedE<-st_transform(breed_season, crs = 4326)
breedE
st_write(breedE, dsn = '~/Dropbox/BGP/COYE/maps', layer = 'COYE.breed_season.sf.WGS84.Ebird', driver = "ESRI Shapefile")

plot(breed_season, col = "#cc503e", border = NA)
plot(breedE, col = "#cc503e", border = NA)

```

```{r}
rng_season <- filter(range_smooth, layer == "range")# %>% 
   # st_geometry()
class(rng_season)
summary(rng_season)
rangeE<-st_transform(rng_season, crs = 4326)
rangeE

st_write(rangeE, dsn = '.', layer = 'COYE.range_smooth.sf.WGS84.Ebird', driver = "ESRI Shapefile")
```

##If reading back in and there is no CRS, give it the CRS assigned and then transform. If title has WGS84, use that projection. IF it doesn't, use mollweide
```{r}
rangeE <- st_read("COYE.range_smooth.sf.WGS84.Ebird.shp") ##when read in the transformed WGS84, it reads in fine. This is what you want for plotting lat/long anyway
class(rangeE)
breedE<-st_read("PABU.breed_season.sf.Ebird.shp") #This one was exported with mollweide
proj4string(rangeE) <- CRS("+proj=moll +lon_0=-90 +x_0=0 +y_0=0 +ellps=WGS84")
rangeE_proj<-st_transform(rangeE, CRS("+proj=moll +lon_0=-90 +x_0=0 +y_0=0 +ellps=WGS84"))

#give it moll projection
mollweide <- "+proj=moll +lon_0=-90 +x_0=0 +y_0=0 +ellps=WGS84"
breedingE <- filter(rangeE_proj,season=="breeding",layer== "range") %>% 
    st_geometry()
plot(breedingE)
class(breedE)
```


